---
title: "DATA 607 Final Project"
author: "James Chun, Pricilla Nakyazze, Haoming Chen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# New York City Property Market

## Overview

This project seeks to research NYC property sales patterns and identify which features may affect sales prices. Such features may include neighborhood, building type, land/gross square footage, tax class and year built. Our goal is to uncover the price distribution in the citywide and trends that could help predict future property values.

Two data sources will be used: a csv file of NYC rolling sales (***citation needed***) and the City of New York API endpoint (Department of Finance). The workflow will include importing data, tidying and cleaning the data, followed by summary statistics, visualizations, analysis and conclusions. One new feature we plan to use is ggmap. This R package perfectly matches our project as it can map NYC property prices to visualize distribution citywide.

Although a database and diagram are not required by the checklist, we still design a MySQL database to store data for its convenience.

## Libraries

```{r message=FALSE}
library(tidyverse)
```

## Import Data

As mentioned before, the main datasets we will be using are NYC rolling sales CSV and City of New York API.

```{r}
nyc_rolling_sales <- read.csv("https://raw.githubusercontent.com/vincent-usny/607-final-/refs/heads/main/NYC_Rolling_Sales.csv")

nyc_api <- jsonlite::fromJSON("https://data.cityofnewyork.us/resource/5ebm-myj7.json?$limit=10000")
```

## Data Cleaning

Preliminary data cleaning is needed for both data sets. To begin, tidying and cleaning will be performed on each dataset seperately.

### NYC Rolling Sales

Begin by removing columns that contain little to no information (i.e. are plagued with NA's). Then, we will also remove columns that contain information irrelevant to our goals. Note that some of the columns state information at present and during time of sale, we will drop the former.

```{r rolling-sales-column-cleaning}
nyc_rolling_sales <- nyc_rolling_sales %>%
  select(-EASE.MENT, -APARTMENT.NUMBER, -BUILDING.CLASS.AT.PRESENT, -SALE.DATE, -RESIDENTIAL.UNITS, -COMMERCIAL.UNITS, -LOT, -BLOCK, -TAX.CLASS.AT.PRESENT, -ZIP.CODE, -ADDRESS) 
```

Next, information in the 'Borough' column will be changed to state the actual name of the borough an not its BBL designation. For reference:<br> 1 - Manhatten<br> 2 - Bronx<br> 3 - Brooklyn<br> 4 - Queens<br> 5 - Staten Island.

```{r rolling-sales-borough-cleaning}
nyc_rolling_sales <- nyc_rolling_sales %>%
  mutate(
    BOROUGH = case_when(
      BOROUGH == 1 ~ "Manhattan",
      BOROUGH == 2 ~ "Bronx",
      BOROUGH == 3 ~ "Brooklyn",
      BOROUGH == 4 ~ "Queens",
      BOROUGH == 5 ~ "Staten Island"
    )
  )
```

The last thing to perform on ***nyc_rolling_sales*** will be removing any rows that contain any missing values in any of the columns. The motivation for this is to drastically cut down on analysis frustrations that may come up later. 

```{r rolling-sales-missing-cleaning}
nyc_rolling_sales <- nyc_rolling_sales %>%
  filter(if_all(everything(), ~ !is.na(.) & . != ""))
```






