---
title: "DATA 607 Week 1 Assignment"
author: "James Chun"
date: "August 30, 2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>

## Introduction
The data presented in this markup was scrapped from the site https://mykbostats.com/players/ and placed into a csv file beforehand. The csv file can be found in the following github repo titled as "KBO_Batters.csv": https://github.com/Ryungje/DATA607/Assignment 1/

I had initially collected this data during a personal project in which I wanted to create a predictive model for KBO baseball player performance. Though the project did not get very far, the data I had collected is still very useful.  

<br>

## Import File from CSV 
This is large data set, as seen from the dimensions.
```{r}
batters <- read.csv("KBO_Batters.csv")
dim(batters)
```
## Create Subset
For our current purposes, we only care about players yearly Team, On-Base Percentage (OBP), Slugging Average (SLG), and OPS. 
```{r}
batters <- batters[c('Name', 'Team', 'Year', 'OBP', 'SLG', 'OPS')]
```

