---
title: "DATA 607 Week 1 Assignment"
author: "James Chun"
date: "August 30, 2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>

## Introduction
The data presented in this markup was scrapped from the site https://mykbostats.com/players/ and placed into a csv file beforehand. The csv file can be found in the following github repo titled as "KBO_Batters.csv": https://github.com/Ryungje/DATA607/tree/main/Assignment%201

I had initially collected this data during a personal project in which I wanted to create a predictive model for KBO baseball player performance. Though the project did not get very far, the data I had collected is still very useful.  

<br>

## Import File from CSV 
This is large data set, as seen from the dimensions.
```{r}
batters <- read.csv("KBO_Batters.csv")
head(batters)
```
## Create Subset
For our current purposes, we only care about players yearly Team, On-Base Percentage (OBP), Slugging Average (SLG), and OPS. 
```{r}
batters <- batters[c('Name', 'Team', 'Year', 'OBP', 'SLG', 'OPS')]
head(batters)
```

## Find Best OPS 
In the simplest way, I shall find the player (disregarding year and team) with the highest OPS.
```{r}
batters[which.max(batters$OPS),]
```

## Conclusion
There are many, and quite complicated, ways to measure a baseball player's performance. The methods presented here are trivial and does not take into account nearly all the data that is available per player. But with that said, the best player in the KBO is Heo Joon.